Exploring influence of REs on models fits

### 2022-05-13 analysis/main/mf-exploration.Rmd
add joint posterior plots

## model without REs; fixed effects only 
f_noREs <- formula(
  target ~ prior * competence * relevance * main_type
)
```

@@

```{r results = 'hide'}
model_SI         <- run_model(f_full,d_critical_clean) 
model_SI_RETitle <- run_model(f_RETitle,d_critical_clean)
model_SI_RESubj  <- run_model(f_RESubj,d_critical_clean)
model_SI_noREs   <- run_model(f_noREs,d_critical_clean)
```

@@

>>>>

## model without REs; fixed effects only 
f_noREs <- formula(
  target ~ prior * competence * relevance * main_type
)

## full model with RE ineractions by-subject
f_full_inters <- formula(
  target ~ prior * competence * relevance * main_type +
    (1 + prior * competence * relevance * main_type || submission_id) +
    (1 | title)
)
```

@@

```{r results = 'hide'}
model_SI         <- run_model(f_full,d_critical_clean) 
model_SI_RETitle <- run_model(f_RETitle,d_critical_clean)
model_SI_RESubj  <- run_model(f_RESubj,d_critical_clean)
model_SI_noREs   <- run_model(f_noREs,d_critical_clean)
model_SI_fullREs <- run_model(f_full_inters, d_critical_clean)
```

@@

## Polina's shenanigans

Check the joint posterior of fixed and random effects, to investigate which of them might cause the pattern of results that we observe. Analysis suggested by Roger.
The following plots shows the joint posterior of fixed and random effects in the xor and some conditions, respectively.

``` {r}
# extract posterior contrast of interest of fixed effects
posteriors <- model_SI_fullREs %>%
  spread_draws(b_prior, `b_prior:main_typexor`, b_relevance, `b_relevance:main_typexor`, b_competence, `b_competence:main_typexor`,
               sd_submission_id__prior, `sd_submission_id__prior:main_typexor`,
               sd_submission_id__Intercept, sd_title__Intercept, 
               sd_submission_id__relevance, `sd_submission_id__relevance:main_typexor`) %>%
    # extract contrasts of interest, especially effect of syntax by-trial
    mutate(prior_xor = b_prior + `b_prior:main_typexor`,
           prior_some = b_prior,
           relevance_xor = b_relevance + `b_relevance:main_typexor`,
           relevance_some = b_relevance,
           competence_xor = b_competence + `b_competence:main_typexor`,
           competence_some = b_competence,
           random_prior_xor = sd_submission_id__prior + `sd_submission_id__prior:main_typexor`,
           random_prior_some = sd_submission_id__prior,
           random_relevance_xor = sd_submission_id__relevance + `sd_submission_id__relevance:main_typexor`,
           random_relevance_some = sd_submission_id__relevance
           ) %>% # subject vs predicate syntax
    select(prior_xor, prior_some, relevance_xor, relevance_some, competence_xor, competence_some,
           random_prior_xor, random_prior_some, sd_submission_id__Intercept, sd_title__Intercept, random_relevance_xor, random_relevance_some) %>%
    gather(key, val)

Helper function for extracting and plotting a heat map of two effects, given a df containing all relevant posterior samples / effects, the effects and a title for the plot. 
```{r}
get_and_plot_2d_posterior <- function(posteriors, fe, re, title) {
  fe_selected <- posteriors %>% filter(key == fe) %>% rename(fe = val)
  re_selected <- posteriors %>% filter(key == re) %>% rename(re = val) %>% select(-key)

  df <- tibble(fe_selected, re_selected)

  p <- ggplot(df, aes(x = fe, y = re)) +
  stat_bin2d(aes(fill = after_stat(density))) +
  ggtitle(title)

  return (p)
}
```

First, we explore how the fixed effects of prior behave against by-title random intercepts and by-subject random slope effects of prior.
```{r}
get_and_plot_2d_posterior(posteriors, "prior_xor", "sd_title__Intercept", "FE of prior for XOR against by-story random intercepts")
#ggsave("FE_prior_RE_story_xor.pdf", width=7, height=5)

get_and_plot_2d_posterior(posteriors, "prior_some", "sd_title__Intercept", "FE of prior for SOME against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "prior_xor", "random_prior_xor", "FE of prior for XOR against by-subj prior RE")

get_and_plot_2d_posterior(posteriors, "prior_some", "random_prior_some", "FE of prior for SOME against by-subj prior RE")
```

Now, double check if there is a "credit assignment" problem for other factors.
```{r}
get_and_plot_2d_posterior(posteriors, "competence_xor", "sd_title__Intercept", "FE of competence for XOR against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "competence_some", "sd_title__Intercept", "FE of competence for SOME against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "relevance_xor", "sd_title__Intercept", "FE of relevance for XOR against by-story random intercepts")

get_and_plot_2d_posterior(posteriors, "relevance_some", "sd_title__Intercept", "FE of relevance for SOME against by-story random intercepts")
```

Explore an xor-only model.
```{r}
f_xor <- formula(
  target ~ prior * competence * relevance +
    (1 + prior + competence + relevance || submission_id) +
    (1 | title)
)
model_xor<- run_model(f_xor, d_critical_clean %>% filter(main_type == "xor")) 
summary(model_xor)
```

```{r}
# get the probability of a negative effect for or
model_xor %>% spread_draws(b_prior) %>%
  select(b_prior) %>%
  gather(key, val) %>%
  group_by(key) %>% mutate(positive = mean(val > 0.05),
                             negative = mean(val < -0.05),
                             no = mean(val %>% between(-0.05, 0.05))) %>%
    summarise(positive_eff = mean(positive),
              no_eff = mean(no),
              negative_eff = mean(negative))
